---
title: "Data extraction and formatting into a table"
format: html
editor: visual
---

# Loading the libraries

```{r}
library(xml2)
library(rvest)
library(stringr)
library(readr)
library(dplyr)
library(here)
```

# Writing the parsing function to get a single dataframe from the multiple html files

```{r}
parse <- function (page) {
  
  # Getting the city name
  
  loc <- page %>% html_elements(xpath = "//h1") %>%
    html_text() %>%
    gsub(pattern=" \\(.*", replacement = "")
  
  # Getting the code
  
  code <- page %>% html_elements(xpath = "//h1") %>%
    html_text() %>%
    gsub(pattern=".*\\(", replacement = "") %>%
    gsub(pattern = "\\)", replacement ="")
  
  # Extracting the user name
  
  user <- page %>% paste( collapse = "\n") %>%
    str_match_all("Par <strong>(.*?)</strong>") %>% 
    as.data.frame() %>% 
    select(2) %>%
    rename(user = X2)
  
  # Getting the ratings
  
  notes <- page %>% html_nodes(".comm") %>%
    html_nodes("table") %>%
    html_nodes("td") %>%
    html_text()
  
  # Getting the title of the rating categories
  
  categories <- page %>% html_elements(xpath = "//th") %>%
    html_text()
  
  categories <- categories[1:9] 
  
  # Getting the dates of the reviews
  
  dates <- page %>% html_nodes(".comm") %>%
    html_text() %>%
    str_extract("([0-9]+(-[0-9]+)+)")
  
  # Getting the positive part of the review
  
  comm_positif <- page %>% html_nodes(".comm") %>%
    html_text() %>%
    gsub(pattern = ".*points positifs : ", replacement =  "") %>%
    gsub(pattern = "Les points négatifs : .*", replacement = "")
  
  # Getting the negative part of the review
  
  comm_negatif <- page %>% html_nodes(".comm") %>%
    html_text() %>%
    gsub(pattern = ".*points négatifs : ", replacement =  "") %>%
    gsub(pattern = "Signaler ce commentaire.*", replacement = "")
  
  # Getting the feedback on the review
  
  feedback <- page %>% html_elements(xpath = "//strong") %>%
    html_text() %>%
    as.numeric() %>%
    na.omit()
  
  # Building the dataframe from the information
  
  data <- tibble(
    code = rep(code, length(comm_positif)),
    loc = rep(loc, length(comm_positif)),
    dates,
    user,
    environment = notes[seq(1,length(notes),9)],
    transports = notes[seq(2,length(notes),9)],
    securite = notes[seq(3,length(notes),9)],
    sante = notes[seq(4,length(notes),9)],
    sp_loisirs = notes[seq(5,length(notes),9)],
    culture = notes[seq(6,length(notes),9)],
    enseignement = notes[seq(7,length(notes),9)],
    commerces = notes[seq(8,length(notes),9)],
    qualite_de_vie = notes[seq(9,length(notes),9)],
    comm_positif,
    comm_negatif,
    agree = feedback[seq(2,length(feedback),3)],
    disagree = feedback[seq(3,length(feedback),3)]
  )
  
  return(data)
}
```

# Applying the function to the html files

## Seting the filepath

```{r}
htmlpath = here::here("Raw/Comments combined/")
```

## Listing all the files from the folder

```{r}
list_pages <- list.files(htmlpath)
```

## Storing all the files into a list

```{r}
pages <- lapply(X = here::here("Raw/Comments combined/", list_pages), 
                FUN = read_html, header=TRUE, sep=",")

head(pages)
```

## Parsing all the html pages and binding it into a single dataframe

```{r}
#| warning: false

commentaires_plm <- purrr::map(pages,parse) %>%
  dplyr::bind_rows()

head(commentaires_plm)
```

# Additionnal formatting of the dataframe

## Dealing with the duplicates

By retaining only the uniquement location/date/user combinations

```{r}

commentaires_plm <- commentaires_plm %>% 
  distinct(loc, dates, user, .keep_all = T)
```

## Formatting the dates column

```{r}
commentaires_plm <- commentaires_plm %>% 
  mutate(dates = lubridate::dmy(commentaires_plm$dates))
```

## Formatting the numeric columns

```{r}
commentaires_plm[5:13] <- lapply(X = commentaires_plm[5:13], 
                                 function(x) as.numeric(x))
```

## Averaging the different categories to get an overall average grade

```{r}
commentaires_plm$average <- commentaires_plm[5:13] %>% 
  rowMeans() %>% 
  round(1) 
```

## Checking the final structure of the dataframe

```{r}
tibble::glimpse(commentaires_plm[, !names(commentaires_plm) %in% "user"])
```

# Exporting the df

```{r}
write_csv(x = commentaires_plm,
          file = here::here("./Derived/comments_PLM_25062026.csv")
          )
```
