---
title: "Descriptive analysis"
format: html
editor: visual
---

# Importing the files and libraries

```{r}
library(dplyr)
library(readr)
library(sf)
library(tmap)
library(ggridges)
library(ggplot2)
library(stringr)
library(here)
```

## Importing the formatted dataset 

Which is derived from the previous script 01-Data_extraction_formatting.qmd 

```{r}
commentaires_plm <- read.csv(file = here::here("Derived/comments_PLM_25062026.csv"))

```

## Importing the spatial administrative layer, for mapping purposes

Which can be downloaded from here: https://geoservices.ign.fr/adminexpress

``` {r}
arr <- sf::st_read(
  here::here("Raw/Arrondissements/ARRONDISSEMENT_MUNICIPAL.shp")
)

```
# Mapping the different indicators by arrondissements

## Checking that the locality names are a suitable key attribute to join the two datasets

```{r}

commentaires_plm$loc <- factor(x = commentaires_plm$loc, 
                               levels = sort(unique(commentaires_plm$loc))
                               )

arr$NOM_M <- factor(x = arr$NOM_M, 
                    levels = sort(unique(arr$NOM_M))
                    )
head(unique(commentaires_plm$loc))

head(unique(arr$NOM_M))

```

### Verifying that no diff exists between the two locality names vectors
``` {r}
setdiff(x = commentaires_plm$loc, 
        y = arr$NOM_M)
```

## Matching the column names with the comments datasets and making the join

``` {r}
arr <- arr %>% rename(loc = NOM_M,
                      pop = POPULATION)

# Joining the geo layer to the comments dataset into a new variable

comments_sf <- merge(x = arr[c("loc", "pop", "geometry")], #retaining the minimal attributes
                     y = commentaires_plm)
```

## Aggregating the different indicators by arrondissement 

Some **NA are generated and that's alright**, because the qualitative variables like "users" and "dates" cannot be grouped, hence turning off the warnings !

``` {r}
#| warning: false

comments_sf_ag <- aggregate(x = comments_sf, 
                            by=list(comments_sf$loc), 
                            FUN= mean) %>%
  mutate(loc = Group.1) %>% # Keeping the arrondissement name in the loc column
  select(-Group.1)
```

## Updating the field to tell in which city each arrondissement is

Which is useful for further grouping by city

``` {r}
# Splitting the location name when a space occurs, to retain only the first bit (i.e. the city name, without the arrondissement part)

comments_sf_ag$ville <- sapply(strsplit(as.character(comments_sf_ag$loc), " "), `[`, 1)
```

## Computing the total of comments by arrondissements

``` {r}

count_comments <- comments_sf$loc %>% 
  table() %>% 
  as_tibble() 

colnames(count_comments) <- c("loc", "nb_comments")

# Joining the result to the table with the aggregate results (groupped by arrondissements)

comments_sf_ag <- merge(comments_sf_ag, count_comments)
```

## Computing the ratio of comments per inhabitants

``` {r}

comments_sf_ag <- comments_sf_ag %>% 
  mutate(comments_ratio = nb_comments/(pop/1000)) #using the population information that was included in the arrondissements spatial layer

# Adding an arrondissement number (to use as labels on the maps ?)

comments_sf_ag <- comments_sf_ag %>% mutate(num_label = as.numeric(str_extract(loc, "\\d+")))

```

# Mapping the various indicators by arrondissements

## Writing the function for the different variables and indicators

``` {r}
generate_gradient_map <- function(var, title) {
  
  # Calculate the average to put it in the map credits
  
  avg_value <- round(mean(comments_sf_ag[[var]], na.rm = TRUE), 2)
  
  # Generate the map
  
  map <- tm_shape(comments_sf_ag) +
    tm_fill( # have to think about this part: rather fixed bins ? 
      col = var,
      palette = "viridis",
      title = title,
      style = 'cont',
      textNA = "",
      legend.is.portrait = FALSE
    ) +
    tm_facets(by = "ville") +
    #tm_labels(text = "num_label") + # labelling to improve
    tm_scalebar(position = c("left", "bottom"), width = 8) +
    tm_layout(
      legend.outside = TRUE,
      legend.title.size = 1,
      legend.title.fontfamily = "Avenir Next",
      legend.title.fontface = "bold",
      legend.text.size = 0.8,
      legend.orientation = "landscape",
      legend.position = tm_pos_out(cell.v = "bottom", cell.h = "center")
    ) +
    tm_credits(
      text = stringr::str_c(
        "Dataset average: ",
        avg_value,
        "\nSource: Comments $ (2012-2026), IGN (2023)."
      ),
      position = tm_pos_out(cell.v = "bottom", cell.h = "center"),
      size = 0.8
    )
  
  # Preparing the filename to export the map
  
  file <- paste("map-", title, ".png", sep = "")
  
  # Saving the map in the figures/ folder
  
  tmap_save(tm = map, 
            filename = here::here("Figures", file), 
            width = 2880,
            height = 1588
            )
}
```

## Mapping the average grading by arrondissements

``` {r}

generate_gradient_map(var = "average", 
                      title = "Average general rating (over 10)")

# Mapping all the other grading categories

categories_to_map <- colnames(comments_sf%>%st_drop_geometry())[6:14]  #can be updated to get a restricted set of categories

categories_to_map # checking the selected categories

for (cat in categories_to_map) {
  category_labels <- c(
    environment = "environment",
    transports = "transportation",
    securite = "safety",
    sante = "health",
    sp_loisirs = "sports and leisure",
    culture = "culture",
    enseignement = "education",
    commerces = "commerce",
    qualite_de_vie = "quality of life"
  )
  cat_label <- category_labels[cat]
  map <- generate_gradient_map(var = cat, 
                        title = paste("Average", cat_label, "rating (over 10)"))
  assign(paste0("map_", cat), map)
}

```
## Mapping the ratio of comments per inhabitants by arrondissements

``` {r}
map_ratio_comments <- generate_gradient_map(var = "comments_ratio", 
                      title = "Number of comments (per 1000 inhabitants)")
```
## Mapping the total count of comments per arrondissements

``` {r}

map_nb_comments <- tm_shape(comments_sf_ag) +
  tm_polygons(
        col = "ville",
        palette = c(
            scales::alpha("red", 0.8),
            scales::alpha("cyan", 0.4),
            scales::alpha("blue", 0.8)
        ),
        legend.show = FALSE,
        style = "cat"
  ) +
  tm_symbols(
    size = "nb_comments",
    title.size = "Number of comments:",
    sizes.legend = c(10, 25, 50, 100, 150),
    sizes.legend.labels = c(10, 25, 50, 100, 150)
  ) +
  tm_facets(by = "ville") +
  tm_layout(
    legend.outside = TRUE,
    legend.title.size = 1,
    legend.title.fontfamily = "Avenir Next",
    legend.title.fontface = "bold",
    legend.text.size = 0.8,
    legend.orientation = "landscape",
    legend.position = tm_pos_out(cell.v = "bottom", cell.h = "center")
  ) +
  tm_scalebar(position = c("left", "bottom"), width = 8) +
  tm_credits(
    text = stringr::str_c(
      "Source: Comments (2012-2026), IGN (2023).\n",
      "Dataset average: ",
      round(mean(comments_sf_ag$average), 2)
    ),
    position = tm_pos_out(cell.v = "bottom", cell.h = "center"),
    size = 0.8
  )
  

tmap_save(
  tm = map_nb_comments,
  filename = here::here("Figures", "map-Number-of-comments.png"), 
  width = 2880,
  height = 1588
  )
```

# Plotting the rating distributions

## Writing the function to generate the plots

``` {r}
generate_ridgeline_plot <- function(var, title) {
 
  ridge_plot <- ggplot(data = commentaires_plm, 
                       aes(x = .data[[var]], 
                           y = arr_label, 
                           fill = ville)
  ) +
    stat_density_ridges(
      geom = "density_ridges_gradient", 
      calc_ecdf = TRUE,
      quantiles = 2, 
      quantile_lines = TRUE
    ) +
    theme_minimal()+
    scale_fill_manual(values = c(
    "PARIS" = scales::alpha("blue", 0.4),
    "LYON" = scales::alpha("red", 0.4),
    "MARSEILLE" = scales::alpha("cyan", 0.4))
    ) +
    # scale_fill_viridis_d(
    #   alpha=.4,
    #   direction = -1,
    #   option = "H"
    #  ) +
    xlab("ratings")+
    scale_x_continuous(breaks = seq(0, 10,by = 1)) +
    theme(legend.position = "none", 
          plot.title = element_text(hjust = 0.5, face = "bold", size=18),
          axis.title.y = element_blank(),
    ) +
    labs(title = title
    ) + 
    facet_wrap(~ville, 
               scales = "free_y")
    
  
  # Displaying the plot
  
  ridge_plot
  
  # Preparing the filename to export the graph
  
  file <- paste("density-ridges-", title, ".png", sep = "")
  
  # Exporting the plot
  
  ggplot2::ggsave(plot = ridge_plot,
                  filename = file,
                  path = here::here("Figures/"),
                  width = 2880,
                  height = 1588,
                  units = "px")
}
```


## Crop the name to just keep the arrondissement number

Used for label on the carts

``` {r}
commentaires_plm$arr_label <- str_extract(commentaires_plm$loc, "\\d{1,2}(ER|E) ARRONDISSEMENT")
```

## Convert it into a factor 

The order is based on the number value so that the labels will appear in the right order when plotting

``` {r}
commentaires_plm <- commentaires_plm %>%
  mutate(
    
    # Extract the number only (e.g. "1" from "1ER ARRONDISSEMENT")
    
    arr_num = as.numeric(str_extract(arr_label, "\\d+")),
    
    # Reorder arr_label factor levels based on [invert order of] arr_num
    
    arr_label = forcats::fct_reorder(arr_label, -arr_num)
  )
```


## Also creating a $ville attribute

That can be used to facet the plots based with respect to the city

``` {r}
commentaires_plm <- commentaires_plm %>%
  mutate(ville = sapply(strsplit(as.character(loc), " "), 
                        function(x) x[1]))
```


## Finally: plotting the distribution by arrondissement in each city, for each indicator

``` {r}

# Plotting the average grade to see that everything works correctly

generate_ridgeline_plot(var = "average", 
                      title = "Average general rating (over 10)")

# Mapping all the other grading categories


for (cat in categories_to_map) {
  category_labels <- c(
    environment = "environment",
    transports = "transportation",
    securite = "safety",
    sante = "health",
    sp_loisirs = "sports and leisure",
    culture = "culture",
    enseignement = "education",
    commerces = "commerce",
    qualite_de_vie = "quality of life"
  )
  cat_label <- category_labels[cat]
  plot <- generate_ridgeline_plot(var = cat, 
                        title = paste("Average", cat_label, "rating (over 10)"))
  assign(paste0("map_", cat), plot)
}

```